%% 摘要

% 中文摘要
\begin{abstract}
    边界表示（Boundary Representation, B-rep）是工业 CAD 内核描述实体几何的标准数据结构，既承载 NURBS 曲面、曲线等连续几何，又记录面—边—顶点之间的拓扑引用关系。如何让神经网络直接产出可被几何内核读取、校验并继续编辑的 B-rep，至今仍是开放问题。已有代表性工作大多绕开几何本身：DeepCAD 让模型生成草图—拉伸等参数化 CAD 操作序列，再交由 CAD 内核执行得到实体；SkexGen 在此基础上把草图与拉伸表征解耦为多组离散码本，并在码本索引上做自回归采样；BrepGen 把面、边、顶点先编码为带拓扑约束的结构化潜变量，再用扩散模型在潜空间中联合采样几何与拓扑。这些路线在中间表征层面分别引入了操作字典、向量量化变分自编码器（VQ-VAE）码本或潜变量分布，使得几何精度上界与拓扑闭合质量同时受到码本规模或潜变量拟合质量的牵制，几何细节的还原能力难以直接通过提升采样分辨率获得。

    针对这一表征瓶颈，本文不再引入显式 VQ-VAE 码本，也不在潜空间上做扩散，而是直接对 B-rep 几何采样点与拓扑面索引做自回归建模，给出了一条点级 NURBS 序列化的端到端 B-rep 生成路线，并完成了与之配套的数据—模型—重建工程闭环。具体工作分三层。数据表征层面，对每个 NURBS 曲面在 UV 参数域上取 $4 \times 4$ 网格共 16 个三维采样点，并拼接 1 个局部面索引，形成长度为 49 的 face\_block；对每条 NURBS 曲线沿弧长方向取 4 个三维点，并在两端拼接邻面索引，形成长度为 14 的 edge\_block。坐标分量先归一化到 $[-1,1]^3$，再经 4000 级标量均匀量化离散化，与 50 个面索引和 3 个结构 token（START / SEP / END）合并为 4053 维统一词表；PAD 在模型侧作为额外填充 token，使模型输出维度增至 4054。模型层面，给出了一种 LLaMA3 风格的标准 next-token 自回归 Transformer，主干由 RoPE、SwiGLU、RMSNorm 与 Grouped Query Attention 组成，序列在主干 RMSNorm 之后直接经 ar\_lm\_head 解码下一 token；代码中保留了 CompoundPointDecoder 作为窗口化分解的可选扩展，但本文主实验固定使用 local\_window\_size = 1，未启用该模块，避免把未经实测的组件写入结论。重建评测层面，基于 OpenCASCADE 实现了从 token 序列经反量化、NURBS 曲面 / 曲线拟合、Sewing 缝合、Solid 构造到 STEP 写出的整段流水线，并在出口处用 BRepCheck\_Analyzer 做内核级合法性校验，从而把"生成结果是否可用"还原为一个可复核的工程指标。

    实验在 DeepCAD 去重子集上完成，训练集 64\,351 条、验证集 6\,096 条。训练阶段使用 4 张 NVIDIA H20 通过 PyTorch DDP 启动 200 个 epoch，端到端耗时 17 小时 28 分钟。主模型 hr2\_patch1 的验证集交叉熵自 Epoch~5 的 0.2357 单调下降到 Epoch~200 的 0.10486，对应困惑度约 1.110，曲线全程未观察到反弹或显著平台期。生成评测以最终 checkpoint 为对象、以 top-k = 1 与 top-p = 0.9 的保守采样配置独立采样 1\,005 条序列，全部成功写出 STEP，BRepCheck 通过 980 条，of saved 与 of attempts 口径下的有效率均为 97.51\%。早期 hr2 基线在同样的采样配置下采样 3\,332 次、写出 STEP 3\,000 个、通过 2\,711 个（of saved 90.40\%，of attempts 81.36\%）。两组结果均明显高于毕业设计任务书"有效率不低于 60\%"的工程红线指标。需要指出的是，两组 run 在采样数量、模型与训练细节上并不严格对齐，本文将其作为工程进展对照而非统计意义上的消融；本文同样未在 ABC 等其他 CAD 数据集上检验跨域泛化能力，也未对拓扑感知排序与 CompoundPointDecoder 进行独立的受控消融，因此不对任何单个组件的贡献量做定量声明。上述未完成的对照实验已在论文末尾以条目形式作为后续工作展开。
    \keywordslist{B-rep, NURBS, 自回归 Transformer, LLaMA3, 点级序列化, OpenCASCADE}
\end{abstract}

% 英文摘要
\begin{engabstract}
    Boundary Representation (B-rep) is the de facto standard for industrial geometric modeling, yet generating B-rep models with complete topology and parametric surfaces directly with deep networks remains an open problem. Recent representative works approach this problem from different angles: DeepCAD generates parametric CAD operation sequences such as sketch-extrude; SkexGen further decouples the sketch and extrude representations and samples on disentangled discrete codebooks with an autoregressive Transformer; BrepGen models B-rep geometry and topology jointly with a diffusion model over a structured latent space. To different extents these methods rely on a VQ-VAE codebook or a latent diffusion prior, and the size of the codebook or the form of the latent distribution directly affects geometric precision and topological closure quality.

    Motivated by the idea of directly modeling B-rep geometry as an autoregressive point sequence without relying on an explicit VQ-VAE codebook, this thesis studies a point-level NURBS serialization framework for end-to-end B-rep generation. The work covers three aspects. (1) A point-level quantization and windowed serialization scheme is designed for B-rep / NURBS data. Each NURBS surface is sampled on its UV parameter domain into 16 three-dimensional points plus one face index (face block of 49 tokens), and each NURBS curve is sampled along arc length into 4 three-dimensional points plus 2 adjacent face indices (edge block of 14 tokens). Every coordinate component is uniformly quantized into 4000 levels, and together with 50 face indices and 3 structural tokens (START / SEP / END), they form a unified vocabulary of 4053 tokens; an additional PAD token is used inside the model so that the model output dimension is 4054. (2) A LLaMA3-style standard next-token autoregressive generative model is implemented, with RoPE, SwiGLU, and RMSNorm as core components. A CompoundPointDecoder is also available in the codebase as an optional window-decomposition extension, but is not activated in the main experiment of this thesis, which uses the standard backbone-norm-plus-ar\_lm\_head next-token head only. (3) An OpenCASCADE-based reverse-reconstruction pipeline is implemented to fit NURBS surfaces and curves from generated tokens, sew them into a Solid, export to STEP, and validate with BRepCheck, providing a self-contained pipeline for measuring generation validity.

    On a DeepCAD subset (64\,351 training / 6\,096 validation samples), the model is trained for 200 epochs with 4 NVIDIA H20 GPUs over 17 hours and 28 minutes. The final model, hr2\_patch1, reaches a validation cross-entropy of 0.10486. On 1\,005 sampled sequences for which a STEP file is successfully written, BRepCheck reports 980 as valid, yielding a valid rate of 97.51\%. For reference, an earlier hr2 baseline (without the patch) attains a valid rate of 90.40\% on 3\,000 written samples. These figures provide preliminary evidence that the proposed point-level serialization and the standard next-token autoregressive pipeline are feasible under the current experimental setting and meet the 60\% valid-rate requirement of the undergraduate thesis brief. No cross-dataset generalization experiment has been performed, and no controlled ablation has been conducted to isolate the independent contribution of the topology-aware ordering or the CompoundPointDecoder; these directions are listed as future work at the end of the thesis.
    \engkeywordslist{B-rep, NURBS, autoregressive Transformer, LLaMA3, point-level serialization, OpenCASCADE}
\end{engabstract}
